Equalizer Programming Guide

### 2008-05-19 Developer/ProgrammingGuide/paper.tex
Finish Programming Guide File Format section, and homogenize file format spec, PG and implementation.
[ ] May break build
[ ] May break existing applications
[ ] Bugfix
[ ] New Feature
[x] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.3.3 for Equalizer 0.5+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add statistics, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-16, 2008 & File Format Appendix [incomplete]\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

The currently used file format is the one-to-one representation of the
server's internal data structures. Its purpose is intermediate, that
is, it will gradually be replaced by automatic resource dedection and
configuration. However the core scalability engine will always use a
similar structure as currently exposed by the file format.

@@

The file format uses the same syntactical structure as VRML. If your
editor supports syntax highlighting and formatting for VRML, you can use
this mode for \textsf{.eqc} files.

@@

The global section defines default values for attributes used by the
individual entities of the server section. The naming convention for
attributes is:

@@

The datatype is one capital letter for the type of the attribute's
value: \textsf{S} for strings, \textsf{C} for a character, \textsf{I}
for an integer and \textsf{F} for floating-point values. Enumation
values are handled as integers. Strings have always to be surrounded by
double quotes '\textsf{''}'. A character has to surrounded by single
quotes '\textsf{'}'.

@@

Global attribute values have useful default parameters, which can be
overriden with an environment variable of the same name. For enumeration
values the corresponding integer value has to be used. The global values
in the config file override these defaults, and are in turn overriden by
the corresponding attributes sections of the associated entities.

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_SATTR\_LAUNCH\_COMMAND\\
  \textbf{Value} & string\\
  \textbf{Default} & ssh -n \%h \%c \textgreater\& \%h.\%n.log [POSIX]\\
                   & ssh -n \%h \%c [WIN32]\\
  \textbf{Details} & The command used by the server to launch nodes
  which could not be connected. The launch command is executed from a
  process forked of the server process. The \% tokens are replaced by the
  server at runtime with concrete data: \%h is replaced by the hostname,
  \%c by the command to launch, including command line arguments and \%n
  by a node-unique identifier. Each command line argument are surrounded
  by launch command quotes.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_CATTR\_LAUNCH\_COMMAND\_QUOTE\\
  \textbf{Value} & character\\
  \textbf{Default} & ' [POSIX]\\
                   & '' [WIN32]\\
  \textbf{Details} & The server uses command line arguments to
  launch render client nodes correctly. This arguments might contain
  white spaces, and therefore have to be surrounded by quotes to
  identify their borders. Certain launch commands or shells use
  different conventions to separate command line arguments.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_TYPE\\
  \textbf{Value} & TCPIP \textbar \ SDP\\
  \textbf{Default} & TCPIP\\
  \textbf{Details} & The protocol for connections. SDP
  programmatically selects the socket direct protocol (AF\_INET\_SDP)
  provided by most InfiniBand protocol stacks.\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_PIPE\_IATTR\_HINT\_THREAD\\
  \textbf{Value} & OFF \textbar \ ON\\
  \textbf{Default} & ON\\
  \textbf{Details} & Determines if all task methods for a pipe and its
  children are executed from a separate operating system thread
  (default) or from the node main thread. Non-threaded pipes have
  certain performance
  limitations and should only be used if really necessary.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STEREO\\
  \textbf{Value} & OFF \textbar \ ON \textbar \ AUTO\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines if the window selects a quad-buffered
  stereo visual. When set to AUTO, the window initialization methods try
  to allocate a stereo visual, but fall back to a mono visual if
  allocation fails.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DOUBLEBUFFER\\
  \textbf{Value} & OFF \textbar \ ON \textbar \ AUTO\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines if the window selects a double-buffered
  stereo visual. When set to AUTO, the window initialization methods try
  to allocate a double-buffered visual, but fall back to a
  single-buffered visual if allocation fails.\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STATISTICS\\
  \textbf{Value} & OFF \textbar \ FASTEST [ON] \textbar \ NICEST\\
  \textbf{Default} & FASTEST [Release Build]\\
                   & NICEST [Debug Build]\\
  \textbf{Details} & Determines how statistics are gathered. OpenGL
  buffers commands, which causes the rendering to be executed at an
  arbitrary point in time. Nicest statistics gathering executes a
  \textsf{Window::finish}, which calls by default \textsf{glFinish}, in
  order to accurately account the rendering operations to the sampled
  task method. However, calling \textsf{glFinish} has a performance
  impact. Therefore, the fastest statistics gathering samples the task
  statistics directly, without finishing the OpenGL commands.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_COLOR\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines the number of color planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of color
  planes specified. AUTO select a visual with a reasonable quality,
  typically eight bits per color.\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_STENCIL\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & UNDEFINED\\
  \textbf{Details} & Determines the number of stencil planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of stencil
  planes specified. By default no stencil planes are requested. Note
  that for depth-compositing and pixel-compositing at least one stencil
  plane is needed.\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_MODE\\
  \textbf{Value} & QUAD \textbar \ ANAGLYPH\\
  \textbf{Default} & QUAD\\
  \textbf{Details} & Selects the algorithm used for stereo
  rendering. QUAD-buffered stereo use the left and right buffers of a
  stereo window (active stereo). Anaglyphics stereo uses
  \textsf{glColorMask} to mask colors for individual eye passes, used in
  conjunction with colored glasses.\\
  \hline
\end{tabularx}\\\vfill

@@

The hostname is the IP address or (resolvable) host name. When used for
the server, the server binds its listening socket only to the specified
name.

A server or node may have multiple connection descriptions.
\clearpage
{\footnotesize\begin{lstlisting}
server
{
    connection // 0-n times, listening connections of the server
    {
        type       TCPIP | SDP
        TCPIP_port unsigned
        hostname   string
    }

@@

For a description of connection attributes please refer to \sref{sGlobal}.

@@

@@

@@

@@

{\footnotesize\begin{lstlisting}
compound
{
    name     string
    channel  string   // where the compound's tasks are executed
    task     [ CLEAR DRAW READBACK ASSEMBLE ] // CULL later
    buffer   [ COLOR DEPTH ]         // default COLOR
    viewport [ viewport ]            // wrt parent compound, sort-first
    range    [ float float ]         // DB-range for sort-last
    pixel    [ int int ]             // pixel decomposition (step size)
    eye      [ CYCLOP LEFT RIGHT ]   // monoscopic or stereo view
    attributes
    {
        stereo_mode                  QUAD | ANAGLYPH    // default QUAD
        // enable visible color components
        stereo_anaglyph_left_mask    [ RED GREEN BLUE ] // default red
        stereo_anaglyph_right_mask   [ RED GREEN BLUE ] // df green blue
        update_FOV                   HORIZONTAL [ON] | VERTICAL | OFF
        assemble                     GPU | CPU          // default GPU
    }

    wall                             // frustum description
    {
        bottom_left  [ float float float ]
        bottom_right [ float float float ]
        top_left     [ float float float ]
    }

@@

    swapbarrier  // compounds with the same barriername sync swap
    {
        name string
    }
    outputframe
    {
        name string
    }
    inputframe
    {
        name string // corresponding output frame
    }
}
\end{lstlisting}}

>>>>

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.3.3 for Equalizer 0.5.1+svn\\[\medskipamount]
  \today
}

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add statistics, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-19, 2008 & File Format Appendix\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

The file format currently used is the one-to-one representation of the
server's internal data structures. Its purpose is intermediate, that is,
it will gradually be replaced by automatic resource dedection and
automatic configuration. However the core scalability engine will
always use a similar structure as currently exposed by the file format.

@@

The file format uses the same syntactical structure as VRML. If your
editor supports syntax highlighting and formatting for VRML, you can use
this mode for editing \textsf{.eqc} files.

@@

The global section defines default values for attributes used by the
individual entities in the server section. The naming convention for
attributes is:

@@

The datatype is one capital letter for the type of the attribute's
value: \textsf{S} for strings, \textsf{C} for a character, \textsf{I}
for an integer and \textsf{F} for floating-point values. Enumation
values are handled as integers. Strings have always to be surrounded by
double quotes '\textsf{''}'. A character has to be surrounded by single
quotes '\textsf{'}'.

@@

Global attribute values have useful default parameters, which can be
overriden with an environment variable of the same name. For enumeration
values the corresponding integer value has to be used. The global values
in the config file override environment variables, and are in turn
overridden by the corresponding attributes sections of the specific
entities.

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_SATTR\_LAUNCH\_COMMAND\\
  \textbf{Value} & string\\
  \textbf{Default} & ssh -n \%h \%c \textgreater\& \%h.\%n.log [POSIX]\\
                   & ssh -n \%h \%c [WIN32]\\
  \textbf{Details} & The command used by the server to launch nodes
  which could not be connected. The launch command is executed from a
  process forked from the server process. The \% tokens are replaced by
  the server at runtime with concrete data: \%h is replaced by the
  hostname, \%c by the command to launch, including command line
  arguments and \%n by a node-unique identifier. Each command line
  argument is surrounded by launch command quotes.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_CATTR\_LAUNCH\_COMMAND\_QUOTE\\
  \textbf{Value} & character\\
  \textbf{Default} & ' [POSIX]\\
                   & '' [WIN32]\\
  \textbf{Details} & The server uses command line arguments to launch
  render client nodes correctly. Certain launch commands or shells use
  different conventions to separate command line arguments. These
  arguments might contain white spaces, and therefore have to be
  surrounded by quotes to
  identify their borders.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_TYPE\\
  \textbf{Value} & TCPIP \textbar \ SDP\\
  \textbf{Default} & TCPIP\\
  \textbf{Details} & The protocol for connections. SDP
  programmatically selects the socket direct protocol (AF\_INET\_SDP)
  provided by most InfiniBand protocol stacks, TCPIP uses normal TCP
  sockets (AF\_INET).\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_PIPE\_IATTR\_HINT\_THREAD\\
  \textbf{Value} & OFF \textbar \ ON\\
  \textbf{Default} & ON\\
  \textbf{Details} & Determines if all task methods for a pipe and its
  children are executed from a separate operating system thread
  (default) or from the node main thread. Non-threaded pipes have
  certain performance limitations and should only be used where necessary.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STEREO\\
  \textbf{Value} & OFF \textbar \ ON \textbar \ AUTO\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines if the window selects a quad-buffered
  stereo visual. When set to AUTO, the window initialization methods try
  to allocate a stereo visual for windows, but fall back to a mono
  visual if allocation fails. For PBuffers, AUTO selects a mono visual.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DOUBLEBUFFER\\
  \textbf{Value} & OFF \textbar \ ON \textbar \ AUTO\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines if the window selects a double-buffered
  stereo visual. When set to AUTO, the window initialization methods try
  to allocate a double-buffered visual for windows, but fall back to a
  single-buffered visual if allocation fails. For PBuffers, AUTO selects
  a single-buffered visual.\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STATISTICS\\
  \textbf{Value} & OFF \textbar \ FASTEST [ON] \textbar \ NICEST\\
  \textbf{Default} & FASTEST [Release Build]\\
                   & NICEST [Debug Build]\\
  \textbf{Details} & Determines how statistics are gathered. OpenGL
  buffers commands, which causes the rendering to be executed at an
  arbitrary point in time. Nicest statistics gathering executes a
  \textsf{Window::finish}, which calls by default \textsf{glFinish}, in
  order to accurately account the rendering operations to the sampled
  task method. However, calling \textsf{glFinish} has a performance
  impact. Therefore, the fastest statistics gathering samples the task
  statistics directly, without finishing the OpenGL commands first. Some
  operations, e.g., frame buffer readback, inherently finish all
  previous OpenGL commands.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_COLOR\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines the number of color planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of color
  planes specified. AUTO selects a visual with a reasonable quality,
  typically eight bits per color.\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_STENCIL\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines the number of stencil planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of stencil
  planes specified. AUTO tries to select a visual with at least one
  stencil plane, but falls back to no stencil planes if allocation
  fails. Note that for depth-compositing and pixel-compositing at least
  one stencil plane is needed.\\
  \hline
\end{tabularx}\\\vfill

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_MODE\\
  \textbf{Value} & QUAD \textbar \ ANAGLYPH\\
  \textbf{Default} & QUAD\\
  \textbf{Details} & Selects the algorithm used for stereo
  rendering. QUAD-buffered stereo uses the left and right buffers of a
  stereo window (active stereo). Anaglyphics stereo uses
  \textsf{glColorMask} to mask colors for individual eye passes, used in
  conjunction with colored glasses.\\
  \hline
\end{tabularx}\\\vfill

@@

The hostname is the IP address or (resolvable) host name. When used for
the server, the server binds its listening socket only to the specified
name.
\clearpage

A server or node may have multiple connection descriptions. A server
binds to all provided connection descriptions. For a node, all
connection descriptions are used while trying to establish a connection
to the node. When auto-launched by the server, all connection
descriptions are passed to the launched node process, which binds to all
provided descriptions.

{\footnotesize\begin{lstlisting}
server
{
    connection // 0-n times, listening connections of the server
    {
        type       TCPIP | SDP
        TCPIP_port unsigned
        hostname   string
    }

@@

A node represents a machine in the cluster, and is one process. It has a
name, a number of connection descriptions and at least one pipe. The
name of the node can be used for debugging, it has no influence on the
execution of Equalizer. For a description of connection attributes
please refer to \sref{sGlobal}.

@@

A pipe represents a graphics card (GPU), and is one execution thread. It
has a name, GPU settings and attributes. The name of a pipe can be used
for debugging, it has no influence on the execution of Equalizer.

The GPU is identified by two parameters, a port and a device. The port
is only used for the GLX window system, and identifies the port number
of the X server, i.e., the number after the colon in the DISPLAY
description (':\textbf{0}.1').

The device identifies the graphics adapter. For the GLX window system
this is the screen number, i.e., the number after the dot in the DISPLAY
description (:0.\textbf{1}). The OpenGL output is always restricted by
glX to the GPU attached to selected screen.

For the AGL window system, the device selects the \textit{n}th display
in the list of online displays. The OpenGL output is optimized for the
selected display, but not restricted to the attached GPU.

For the WGL window system, the device selects the \textit{n}th GPU in
the system. The GPU can be offline, in this case only PBuffer windows
can be used. In order to restrict the OpenGL output to the GPU, the
\textsf{WGL\_NV\_gpu\_affinity} extension is used. If the extension is
not present, the window is opened on the \textit{n}th monitor, but
OpenGL commands are sent to all GPU's.

The viewport of the pipe can be used to override the pipe
resolution. The viewport is defined in pixels. The x and y parameter of
the viewport are currently ignored. The default viewport is
automatically detected. For offline GPU's, a default of 4096x4096 is
used.

For a description of pipe attributes please refer to \sref{sGlobal}.

@@

A window represents an OpenGL drawable and holds an OpenGL context. It
has a name, a viewport and attributes. The name of a window can be used
for debugging, it has no influence on the execution of Equalizer.

The viewport of the window is relative to the pipe. It can be specified
in relative or absolute coordinates. Relative coordinates are normalized
coordinates with respect to the pipe, e.g., a viewport of \textsf{[ 0.25
  0.25 0.5 0.5 ]} creates a window in the middle of the screen, using
50\% of the pipe's size. Absolute coordinates are integer pixel values,
e.g., a viewport of \textsf{[ 50 50 800 600 ]} creates a window 50 pixels
from the upper-left corner, sized 800x600 pixels, regardless of the
pipe's size. The default viewport is \textsf{[ 0 0 1 1 ]}, i.e., a
full-screen window.

For a description of window attributes please refer to \sref{sGlobal}.

@@

A channel is a two-dimensional area within a window. It has a name,
viewport and attributes. The name of the channel is used to identify the
channel in the respective compounds. It should be unique within the config.

The viewport of the channel is relative to the window. As for windows,
it can be specified in relative or absolute coordinates. The default
viewport is \textsf{[ 0 0 1 1 ]}, i.e., fully covering its window.

For a description of channel attributes please refer to \sref{sGlobal}.

@@

Compounds are the basic datastructure describing the rendering
setup. They use channels for rendering. Please refer to
\sref{sCompounds} for a description of compound operation logics.

The name of the compound is used for the default names of swap barriers
and output frames. The channel references the first channel of the same
name in the configuration.

Compound tasks describe the operations the compound executes. The
default is all tasks for compounds with no children (leaf compounds) and
\textsf{READBACK ASSEMBLE} for all others. The readback and assemble
tasks are only executed if the compound has output frames or input
frames, respectively. Tasks are not inherited by the children of a
compound.

The buffer defines the default frame buffer attachments read back by
output frames. The viewport restricts the rendering to the area relative
to the parent compound. The range restricts the database range, relative
to the parent. The pixel setting selects the pixel decomposition
parameters, relative to the parent. The eye pass select the mono, left
or right eye passes rendered by the compound. All these attributes are
inherited by the children of a compound. Viewport, range and pixel
parameters are cumulative.

For a description of compound attributes please refer to \sref{sGlobal}.

A wall or projection description is used to define the view frustum of
the compound. The frustum is inherited and typically only defined on the
top-most compound. The last specified frustum description is used. Sizes
are specified in meters. \fig{fFrustra} illustrates the frustum
parameters.

A swap barrier is used to synchronize the output of multiple
compounds. All swap barriers of the same name are synchronized The
default name is \textsf{barrier[.root\-Com\-pound\-Name]}.

Output frames transport frame buffer contents to input frames of the
same name. If the compound has a name, the default frame name is
\textsf{frame.compoundName}, otherwise the default name is
\textsf{frame.channelName}. The frame buffer attachments to read back
are inherited from the compound, but can be overridden by output frames.

{\footnotesize\begin{lstlisting}
compound
{
    name     string
    channel  string   // where the compound's tasks are executed
    task     [ CLEAR DRAW READBACK ASSEMBLE ] // CULL later
    buffer   [ COLOR DEPTH ]         // default COLOR
    viewport [ viewport ]            // wrt parent compound, sort-first
    range    [ float float ]         // DB-range for sort-last
    pixel    [ int int ]             // pixel decomposition (step size)
    eye      [ CYCLOP LEFT RIGHT ]   // monoscopic or stereo view
    attributes
    {
        stereo_mode                  QUAD | ANAGLYPH    // default QUAD
        stereo_anaglyph_left_mask    [ RED GREEN BLUE ] // default red
        stereo_anaglyph_right_mask   [ RED GREEN BLUE ] // df green blue
        update_FOV                   HORIZONTAL [ON] | VERTICAL | OFF
    }

    wall       // frustum description
    {
        bottom_left  [ float float float ]
        bottom_right [ float float float ]
        top_left     [ float float float ]
    }
    projection // alternate frustum description, last one wins
    {
        origin       [ float float float ]
        distance     float
        fov          [ float float ]
        hpr          [ float float float ]
    }

@@

    swapbarrier  // compounds with the same barriername sync swap
    {
        name string
    }
    outputframe
    {
        name   string
        buffer [ COLOR DEPTH ]
    }
    inputframe
    {
        name string // corresponding output frame
    }
}
\end{lstlisting}}
